Aracna: An Open-Source Quadruped Platform for Evolutionary Robotics

### 2012-06-01 paper_aimsa/conclusion.tex
Added splines example figure, wrapped lines of text for easier merging

\section{Conclusion}

Conclusion and Future Work
We have presented a new reinforcement-learning-based algorithm for optimizing a quadrupedal gait for linear speed. We implemented and tested six learning strategies for parameterized gaits and compared them to gaits produced by neural networks evolved with the HyperNEAT generative encoding. Though both methods resulted in an improvement over the robot’s previous naïve gaits, based on the statistics collected, RL_PoWER has a more elegant and consistent performance.

Based on previous studies on evolving gait on quadruped robots[xx], RL_PoWER produced higher performing gaits than all of the parameterized methods. Its best-performing gait traveled XX body lengths per minute.  HyperNEAT, showed in the previous studies, achieved the competitive best performance with RL_PoWER. HyperNEAT gaits

Though over 900 trials have been made to investigate the applicability of RL_PoWER to quadruped robots. It is difficult to gather the enough trials that would be necessary to properly rank the methods. One direction for future work could be to obtain many more trials. But due to the physical limitations, obtaining  One solution to this is simulation. Because of the low cost of simulation, it would produce the necessary volume of trials to allow the learning methods to be effective, and the hardware trials would serve to continuously ground and refine the simulator. [KYRREG XX]

>>>>

\section{Conclusion and Future Work}

We have presented a new reinforcement-learning-based algorithm for
optimizing a quadrupedal gait for linear speed. We implemented and
tested six learning strategies for parameterized gaits and compared
them to gaits produced by neural networks evolved with the HyperNEAT
generative encoding. Though both methods resulted in an improvement
over the robot’s previous \naive gaits, based on the statistics
collected, RL\_PoWER has a more elegant and consistent performance.

Based on previous studies on evolving gait on quadruped robots[xx],
RL\_PoWER produced higher performing gaits than all of the
parameterized methods. Its best-performing gait traveled XX body
lengths per minute.  HyperNEAT, showed in the previous studies,
achieved the competitive best performance with RL\_PoWER. HyperNEAT
gaits

Though over 900 trials have been made to investigate the applicability
of RL\_PoWER to quadruped robots. It is difficult to gather the enough
trials that would be necessary to properly rank the methods. One
direction for future work could be to obtain many more trials. But due
to the physical limitations, obtaining One solution to this is
simulation. Because of the low cost of simulation, it would produce
the necessary volume of trials to allow the learning methods to be
effective, and the hardware trials would serve to continuously ground
and refine the simulator. [KYRREG XX]

### 2012-06-01 paper_aimsa/introduction.tex
Added splines example figure, wrapped lines of text for easier merging

What the problem is...

@@

General learning:Learning gaits Various learning algorithms have been tested to be effective for legged robots. Their studies tested the viability of applying RL method to legged robot gaits learning. Competitive performance has been verified on methods such as HyperNEAT (Yosinski et al, 2011) and GA(Chernova et al, 2004) and other  methods(Hornby et al., 2005; Zykov et al., 2004; T´ellez et al., 2006; Valsalam and Miikkulainen, 2008). Despite the excellent performance of these algorithms, a major part of achieving good performance relies in tuning the parameters for the evolutionary algorithms(Kormushev et al, 2006). Here we present a different way for learning gaits using RL(Hill climbing). In our experiment, it turned out this method achieved competitive performance with Problem Definition:The gait-learning problem is defined to find a gait that maximizes some specific metric. Mathematically, we define a gait as a function that specifies a vector of commanded motor positions for a robot over time. Gaits without feedback --- also called open-loop gaits --- can be defined as \be\vec{x} = g(t)\ee According to the definition, open-loop gaits are deterministic. One particular gait should behave exactly the same when it is run from trial to trial. However, the actual robot motion and fitness measured will vary due to the noisiness of the real world physics.In our trials, one gait pattern was generated and sent to the robot and executed in an open loop manner, as defined. In particular, in this paper, we chose to compare open-loop gaits generated specifically by the RL\_POWER and HyperNEAT. The particular metric used will be described later.

>>>>

@@

Various learning algorithms have been tested to be effective for
legged robots. Their studies tested the viability of applying RL
method to legged robot gaits learning. Competitive performance has
been verified on methods such as HyperNEAT (Yosinski et al, 2011) and
GA(Chernova et al, 2004) and other methods(Hornby et al., 2005; Zykov
et al., 2004; T´ellez et al., 2006; Valsalam and Miikkulainen,
2008). Despite the excellent performance of these algorithms, a major
part of achieving good performance relies in tuning the parameters for
the evolutionary algorithms(Kormushev et al, 2006). Here we present a
different way for learning gaits using RL(Hill climbing). In our
experiment, it turned out this method achieved competitive performance
with Problem Definition:The gait-learning problem is defined to find a
gait that maximizes some specific metric. Mathematically, we define a
gait as a function that specifies a vector of commanded motor
positions for a robot over time. Gaits without feedback --- also
called open-loop gaits --- can be defined as

\be\vec{x} = g(t)\ee

According to the definition, open-loop gaits are deterministic. One
particular gait should behave exactly the same when it is run from
trial to trial. However, the actual robot motion and fitness measured
will vary due to the noisiness of the real world physics.In our
trials, one gait pattern was generated and sent to the robot and
executed in an open loop manner, as defined. In particular, in this
paper, we chose to compare open-loop gaits generated specifically by
the RL\_POWER and HyperNEAT. The particular metric used will be
described later.

### 2012-06-01 paper_aimsa/methods.tex
Added splines example figure, wrapped lines of text for easier merging

how we tried to solve it...

Parameterized Gaits by RL\_PoWER: Here we used an RL approach to change the complexity of the policy representation dynamically while the trial is running. In Petar et al’ s studies on reducing energy consumption for bipedal robots [xx], a mechanism that can evolve the policy parameterization was used. The method starts from a very simple parameterization and gradually increases its representational capability. The method tested to be able to generate an adaptive policy parameterization that can accommodate increasingly more complex policies. In the work described in [xx], the policy generated by this approach can reach the global optimum at a faster rate. Also found is that the chance of converging to a suboptimal solution is reduced, because in the lower-dimensional representation the effect is less exhibited. Spline policy representationThe simplest model with back-compatibility is geometric splines. For a given model f(x) with K knots, we can preserve the exact shape of the generated curve while adding extra knots to the original spline. Say, if we put one additional knot between every two consecutive knots of the original spline, we end up with a 2K - 1 knots and a spline that has the same shape as the original one. In order to do this, we need to define an algorithm for evolving the parameterization from K to L knots (L > K ), which is formulated in Algorithm 1.  Without loss of generality, the policy parameters are normalized into [0, 1], and appropriately scaled/shifted as necessary later upon use.[pic]Fig. 2 illustrates the process of using spline representation for the evolving policy parameterization. Fig. 1 shows an example for a reinforcement learning process using evolving policy parameterization to approximate an unknown function. Petar et alIntegrating evolving policy parameterization with Hill-climbing(RL?)Kober et al  proposed (Kober et al) a RL algorithm called Policy learning by Weighting Exploration with the Returns(PoWER), which is based on Expectation-Maximization  algorithm (EM). Our proposed technique for evolving the policy parameterization are combined with this EM-based RL algorithm PoWER [xx], due to its relatively fewer parameters that need tuning. Also because PoWER demonstrated high performance in tasks learned directly on real robots, such as pendulum swing-up and pancake flipping task [9].We evolved the policy parameterization only on those past trials ranked the highest by the importance sampling technique used by the PoWER algorithm(). The intuition behind is that high-ranked parameterizations have more potential to perform even better in the future. Besides, evolving all the parameterizations requires significantly more efforts. (Bongard et al.,2006) Since our experiment is done on physical robot, explore all the variations of every parameterization is not practical. A promising extension can be, using physical simulator for the robot, to further explore the parameterization space by evolving more parameterizations. Gaits of RL\_POWER:For the experiment, we have 3 knots for each spline and 8 splines in total for the servos. The servo in the hip is not used in our experiment. Previous work has verified that quadruped gaits perform better when they are coordinated (Clune et al., 2009a, 2011; Valsalam and Miikkulainen, 2008). So the gaits For each set of splines, we calculate its corresponding parameterized gait within a time cycle. Given the pattern of one unit cycle, then apply the same pattern to every cycle throughout the whole period of trial. Specifically, each spline is stretched to servo positions as following:   rescale for one cycle: 	Unit Cycle :s(si).cyclePos = s(si).YY .* posRange + posMin;    		s(si).cyclePos = s(si).cyclePos(:);    	Repeat cycle,      	Cycle[i] = UnitCycle% the last pos in a cycle is the same as the first pos in the next cycle, from [pic for one gait]

>>>>

\subsection{Parameterized Gaits by RL\_PoWER}

Here we used an RL approach to change the complexity of the policy
representation dynamically while the trial is running. In Petar et al’
s studies on reducing energy consumption for bipedal robots [xx], a
mechanism that can evolve the policy parameterization was used. The
method starts from a very simple parameterization and gradually
increases its representational capability. The method tested to be
able to generate an adaptive policy parameterization that can
accommodate increasingly more complex policies. In the work described
in [xx], the policy generated by this approach can reach the global
optimum at a faster rate. Also found is that the chance of converging
to a suboptimal solution is reduced, because in the lower-dimensional
representation the effect is less exhibited.

\subsection{Spline policy representation}

The simplest model with back-compatibility is geometric
splines. For a given model f(x) with K knots, we can preserve the
exact shape of the generated curve while adding extra knots to the
original spline. Say, if we put one additional knot between every two
consecutive knots of the original spline, we end up with a 2K - 1
knots and a spline that has the same shape as the original one. In
order to do this, we need to define an algorithm for evolving the
parameterization from $K$ to $L$ knots ($L > K$), which is formulated in
Algorithm \edit{proper reference}.  Without loss of generality, the policy parameters are
normalized into $[0, 1]$, and appropriately scaled/shifted as necessary
later upon use.

\figp{powerSplinesExample}{.75}{Example figure... from Petar's paper...}

[pic]Fig. 2 illustrates the process of using spline
representation for the evolving policy parameterization. Fig. 1 shows
an example for a reinforcement learning process using evolving policy
parameterization to approximate an unknown function. Petar et
alIntegrating evolving policy parameterization with
Hill-climbing(RL?)Kober et al proposed (Kober et al) a RL algorithm
called Policy learning by Weighting Exploration with the
Returns(PoWER), which is based on Expectation-Maximization algorithm
(EM). Our proposed technique for evolving the policy parameterization
are combined with this EM-based RL algorithm PoWER [xx], due to its
relatively fewer parameters that need tuning. Also because PoWER
demonstrated high performance in tasks learned directly on real
robots, such as pendulum swing-up and pancake flipping task [9].We
evolved the policy parameterization only on those past trials ranked
the highest by the importance sampling technique used by the PoWER
algorithm(). The intuition behind is that high-ranked
parameterizations have more potential to perform even better in the
future. Besides, evolving all the parameterizations requires
significantly more efforts. (Bongard et al.,2006) Since our experiment
is done on physical robot, explore all the variations of every
parameterization is not practical. A promising extension can be, using
physical simulator for the robot, to further explore the
parameterization space by evolving more parameterizations. Gaits of
RL\_POWER:For the experiment, we have 3 knots for each spline and 8
splines in total for the servos. The servo in the hip is not used in
our experiment. Previous work has verified that quadruped gaits
perform better when they are coordinated (Clune et al., 2009a, 2011;
Valsalam and Miikkulainen, 2008). So the gaits For each set of
splines, we calculate its corresponding parameterized gait within a
time cycle. Given the pattern of one unit cycle, then apply the same
pattern to every cycle throughout the whole period of
trial. Specifically, each spline is stretched to servo positions as
following: rescale for one cycle: Unit Cycle :s(si).cyclePos =
s(si).YY .* posRange + posMin; s(si).cyclePos = s(si).cyclePos(:);
Repeat cycle, Cycle[i] = UnitCycle \% the last pos in a cycle is the
same as the first pos in the next cycle, from [pic for one gait]

### 2012-06-01 paper_aimsa/results.tex
Added splines example figure, wrapped lines of text for easier merging

RL_POWER Gaits
The results for the gaits evolved by RL_POWER are shown
in Figure 8 and Table 2. A total of 900 evaluations were performed for HyperNEAT (300 in each of three runs). Overall the RL_POWER gaits were the faster and more stable by far, beating the HyperNEAT when comparing either average or best gaits

(pic)RL_POWER AND HyperNEAT performance chart

or best gaits. We believe that this is because RL_POWER uses evolvable splines to represent gaits As explained earlier,  
The single best gait found during this study had a speed of
XXXX body lengths/minute, XX%
better than the best HyperNEAT gait. Figure 9 shows a typical RL_POWER gait that had high fitness. The pattern of motion is less complex and more regular than the HyperNEAT, in that patterns of multiple motors are coordinated.
A corresponding observation of this property is that the noisiness of RL_POWER is significantly lower than that of HyperNEAT. One reason of this is that the when Expectation Maximization methodology is used, explorations happened in the vicinity space of the highest ranked patterns. Even some individual HyperNEAT gaits turned out very effective, the average noisiness of HyperNEAT is very large. For example, as found in [xx], the mean of HyperNEAT generation-champion’s performance was XX body lengths/minute (+-XX SD), but it had a max of 38 and a min of 3. RL_POWER, on the other hand,  proved to be stable and much less noisy throughout the entire experiment. The maximum SD over all three runs is +- XX SD. This may also be due to a much larger space of possibilities of HyperNEAT than the more heuristically guided reinforcement learning.

>>>>

RL\_POWER Gaits

The results for the gaits evolved by RL\_POWER are shown in Figure 8
and Table 2. A total of 900 evaluations were performed for HyperNEAT
(300 in each of three runs). Overall the RL\_POWER gaits were the
faster and more stable by far, beating the HyperNEAT when comparing
either average or best gaits

(pic)RL\_POWER AND HyperNEAT performance chart

or best gaits. We believe that this is because RL\_POWER uses evolvable splines to represent gaits As explained earlier,  
The single best gait found during this study had a speed of

XXXX body lengths/minute, XX\%

better than the best HyperNEAT gait. Figure 9 shows a typical
RL\_POWER gait that had high fitness. The pattern of motion is less
complex and more regular than the HyperNEAT, in that patterns of
multiple motors are coordinated.

A corresponding observation of this property is that the noisiness of
RL\_POWER is significantly lower than that of HyperNEAT. One reason of
this is that the when Expectation Maximization methodology is used,
explorations happened in the vicinity space of the highest ranked
patterns. Even some individual HyperNEAT gaits turned out very
effective, the average noisiness of HyperNEAT is very large. For
example, as found in [xx], the mean of HyperNEAT generation-champion’s
performance was XX body lengths/minute (+-XX SD), but it had a max of
38 and a min of 3. RL\_POWER, on the other hand, proved to be stable
and much less noisy throughout the entire experiment. The maximum SD
over all three runs is +- XX SD. This may also be due to a much larger
space of possibilities of HyperNEAT than the more heuristically guided
reinforcement learning.
